Contextual Combinatorial Cascading Bandits

### 2015-11-29 contextual combinatorial cascading bandits.tex
update part of 'or' case

To prove this theorem, we could start from estimating $\EE[\Delta_{t, \bA_t}|\cH_t]$ for every time $t$. In fact, we can derive the following lemma.
\begin{lemma}
\label{lem:DeltaEstimate}
For any time $t$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$ where $A = (a_1, \ldots, a_{\abs{A}})$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k' \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
$$
\end{lemma}

@@

\begin{lemma}
\label{lem:increasing} 
Given parameters $0= \gamma_0 \leq \gamma_1 \leq ... \leq \gamma_{K-1} \leq 1$, $f(A, w)$ is increasing with respect to $w$; equivalently, suppose $0 \leq w \leq w' \leq 1$ as column vectors in $\RR^L$, we have for any $A\in\cS$,
$$
f(A, w) \leq f(A, w').
$$
\end{lemma}
\begin{proof}
It suffices to prove that $\cA = [m]$ where $1 \leq m \leq K$. Because for $1 \leq k \leq m$,
\begin{align*}
&~~-\gamma_{k-1} + \sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i') + \prod_{j=k+1}^{m}w_j'\\
&\geq \gamma_{k-1} (-1 + \sum_{i=k+1}^m (\prod_{j=k+1}^{i-1}w_j')(1- w_i') +  \prod_{j=k+1}^{m}w_j')\\
&=\gamma_{k-1} \cdot 0 = 0,
\end{align*}
then
\begin{align*}
&~~\gamma_{k-1} (1 - w_k) + w_k\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i')\\
&\qquad \qquad + w_k\prod_{j=k+1}^{m}w_j'\\
& \leq \gamma_{k-1} (1 - w_k') + w_k'\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i')\\
&\qquad \qquad + w_k'\prod_{j=k+1}^{m}w_j'.
\end{align*}
Therefore, 
\begin{align*}
& f(A; w_1,...,w_{k-1},w_k,w_{k+1}',...,w_m')\\
&=\sum_{i=2}^{k-1} \gamma_{i-1} (\prod_{j=1}^{i-1}w_j)(1 - w_i) + (\prod_{j=1}^{k-1}w_j) [\gamma_{k-1} (1 - w_k) \\
&\qquad + w_k\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i') + w_k\prod_{j=k+1}^{m}w_j']\\
&\leq \sum_{i=2}^{k-1} \gamma_{i-1} (\prod_{j=1}^{i-1}w_j) (1 - w_i) + (\prod_{j=1}^{k-1}w_j) [\gamma_{k-1} (1 - w_k') \\
&\qquad + w_k'\sum_{i=k+1}^m \gamma_{i-1} (\prod_{j=k+1}^{i-1}w_j')(1- w_i') + w_k'\prod_{j=k+1}^{m}w_j']\\
&=f(A; w_1,...,w_{k-1},w_{k}',w_{k+1}',...,w_m').
\end{align*}
Then we obtain the result.
\end{proof}

\begin{lemma}
\label{lem:estimateTech}
Suppose $0 \leq p_1,...,p_m \leq 1$ and $u_1,...,u_m \geq 0$. Let $w_i = \min\{p_i + u_i, 1\}, 1 \leq i \leq m$. Then
$$
f(A^{(m)}, w) \leq f(A^{(m)}, p) + \sum_{k=1}^{m} \gamma_k' u_k,
$$
where $A^{(m)} = (1,2,\ldots,m), w = (w_1, \ldots, w_m)$ and $p = (p_1, \ldots, p_m)$.
\end{lemma}
\begin{proof}
We prove this by induction. It holds obviously for $m=1$.
\begin{align*}
&~~f(m+1, w)\\
&= f(m, w) -\prod_{k=1}^{m} w_k + \gamma_{m}  (\prod_{k=1}^{K} w_k) (1 - w_{K+1})+ \prod_{k=1}^{K+1} w_k\\
&= f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} w_k) (1 - w_{m+1})\\
&\leq f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} p_k) (1 - w_{m+1})\\
&\leq f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} p_k)  (1 - p_{m+1} - u_{m+1})\\
&\leq f(m, p) +  \sum_{k=1}^{m} \gamma_k' u_k + (\gamma_{m} - 1) (1 - p_{m+1}) \prod_{k=1}^{m} p_k\\
&~~ + (1 - \gamma_{m}) u_{m+1}\\
&\leq f(m+1, p) + \sum_{k=1}^{m+1} \gamma_k' u_k
\end{align*}
\end{proof}

@@

\begin{lemma}
Suppose $A = (a_1, ..., a_{\abs{A}})$. Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. Then we have the following properties.
\begin{itemize}
\item[(1)] $f(A, w) \leq p_A + \gamma_{\abs{A} - 1}$;
\item[(2)] $f(B_{k+1}, w) \leq f(B_k, w)$;
\item[(3)] $p_{B_{k+1}} \leq p_{B_k}$;
\item[(4)] $f(B_k, w) \leq p_{B_{k+1}} + \gamma_{k}$.
\end{itemize}	
\end{lemma}

>>>>

Suppose the reward function is disjunctive. At every time $t$, the forecaster recommends a list $\bA_t$. Then the user checks from the first item to the end. The weight $\bw_t(a)$ denote whether the user click on item $a$ or not. The user stops if she clicks on one item of $\bA_t$ or she has checked all items of $\bA_t$. If the user clicks on the first item, then the learning agent receives reward $1$. If the user clicks on the $k$-th reward of $\bA_t$, then the reward is $\gamma_k$. If the user doesn't click on any item, then the reward is $0$. Here we further assume $\gamma_2 \geq \gamma_3 \geq \ldots \geq \gamma_K$. If we define a function $f_{\vee}$ on $A = (a_1, \ldots, a_{\abs{A}}) \in \cS$ and $w = (w(1), \ldots, w(L))$ by
\begin{align*}
&f_{\vee} : \cS \times [0,1]^E \to [0,1]\\
&f_{\vee}(A,w) = \sum_{k=1}^{\abs{A}}\gamma_k (\prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k).
\end{align*}
Here we use notation that $\gamma_1 = 1$. Then the expected reward of $\bA_t$ is $f_{\vee}(\bA_t, \theta_*^{\top}x_t)$.

Then we have a relationship between $f$ and $f_{\vee}$
\begin{equation}
\label{eq:relation}
\begin{split}
&f_{\vee}(A, w; 1, \gamma_2, \ldots, \gamma_K) \\
&~~= 1 - f(A, 1 - w; 0, 1-\gamma_2, \ldots, 1-\gamma_K)
\end{split}
\end{equation}
If $\gamma$ is decreasing, then $1 - \gamma$ is increasing. Thus we have $1 - \gamma$ a feasible parameters for $f$. Then we could have a similar result for disjunctive case.
\begin{theorem}
Disjunctive case
\end{theorem}

To prove theorem \ref{thm:main}, we could start from estimating $\EE[\Delta_{t, \bA_t}|\cH_t]$ for every time $t$. In fact, we can derive the following lemma.
\begin{lemma}
\label{lem:DeltaEstimate}
For any time $t$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$ where $A = (a_1, \ldots, a_{\abs{A}})$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k' \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
$$
If $f_{\vee}(A_t^*, \bU_t) \leq f_{\vee}(A, \bU_t)$ where $A = (a_1, \ldots, a_{\abs{A}})$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
$$
\end{lemma}

@@

\begin{lemma}
\label{lem:increasing} 
Given parameters $0= \gamma_0 \leq \gamma_1 \leq ... \leq \gamma_{K-1} \leq 1$, $f(A, w)$ is increasing with respect to $w$; equivalently, suppose $0 \leq w \leq w' \leq 1$ as column vectors in $\RR^L$, we have for any $A\in\cS$,
$$
f(A, w) \leq f(A, w').
$$
By relation (\ref{eq:relation}), it also holds that
$$
f_{\vee}(A, w) \leq f_{\vee}(A, w').
$$
\end{lemma}
\begin{proof}
It suffices to prove that $\cA = [m]$ where $1 \leq m \leq K$. Because for $1 \leq k \leq m$,
\begin{align*}
&~~-\gamma_{k-1} + \sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i') + \prod_{j=k+1}^{m}w_j'\\
&\geq \gamma_{k-1} (-1 + \sum_{i=k+1}^m (\prod_{j=k+1}^{i-1}w_j')(1- w_i') +  \prod_{j=k+1}^{m}w_j')\\
&=\gamma_{k-1} \cdot 0 = 0,
\end{align*}
then
\begin{align*}
&~~\gamma_{k-1} (1 - w_k) + w_k\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i')\\
&\qquad \qquad + w_k\prod_{j=k+1}^{m}w_j'\\
& \leq \gamma_{k-1} (1 - w_k') + w_k'\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i')\\
&\qquad \qquad + w_k'\prod_{j=k+1}^{m}w_j'.
\end{align*}
Therefore, 
\begin{align*}
& f(A; w_1,...,w_{k-1},w_k,w_{k+1}',...,w_m')\\
&=\sum_{i=2}^{k-1} \gamma_{i-1} (\prod_{j=1}^{i-1}w_j)(1 - w_i) + (\prod_{j=1}^{k-1}w_j) [\gamma_{k-1} (1 - w_k) \\
&\qquad + w_k\sum_{i=k+1}^m \gamma_{i-1}(\prod_{j=k+1}^{i-1}w_j')(1- w_i') + w_k\prod_{j=k+1}^{m}w_j']\\
&\leq \sum_{i=2}^{k-1} \gamma_{i-1} (\prod_{j=1}^{i-1}w_j) (1 - w_i) + (\prod_{j=1}^{k-1}w_j) [\gamma_{k-1} (1 - w_k') \\
&\qquad + w_k'\sum_{i=k+1}^m \gamma_{i-1} (\prod_{j=k+1}^{i-1}w_j')(1- w_i') + w_k'\prod_{j=k+1}^{m}w_j']\\
&=f(A; w_1,...,w_{k-1},w_{k}',w_{k+1}',...,w_m').
\end{align*}
Then we obtain the result.
\end{proof}

\begin{lemma}
\label{lem:estimateTech}
Suppose $0 \leq p_1,...,p_m \leq 1$ and $u_1,...,u_m \geq 0$. Let $w_i = \min\{p_i + u_i, 1\}, 1 \leq i \leq m$. Then
$$
f(A^{(m)}, w) \leq f(A^{(m)}, p) + \sum_{k=1}^{m} \gamma_k' u_k,
$$
where $A^{(m)} = (1,2,\ldots,m), w = (w_1, \ldots, w_m)$ and $p = (p_1, \ldots, p_m)$, $\gamma$ is increasing. It also holds that
$$
f_{\vee}(A^{(m)}, w) \leq f_{\vee}(A^{(m)}, p) + \sum_{k=1}^m \gamma_k u_k,
$$
for any positive sequence $\gamma$.
\end{lemma}
\begin{proof}
We prove this by induction. It holds obviously for $m=1$.
\begin{align*}
&~~f(m+1, w)\\
&= f(m, w) -\prod_{k=1}^{m} w_k + \gamma_{m}  (\prod_{k=1}^{K} w_k) (1 - w_{K+1})+ \prod_{k=1}^{K+1} w_k\\
&= f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} w_k) (1 - w_{m+1})\\
&\leq f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} p_k) (1 - w_{m+1})\\
&\leq f(m, w) + (\gamma_{m} - 1) (\prod_{k=1}^{m} p_k)  (1 - p_{m+1} - u_{m+1})\\
&\leq f(m, p) +  \sum_{k=1}^{m} \gamma_k' u_k + (\gamma_{m} - 1) (1 - p_{m+1}) \prod_{k=1}^{m} p_k\\
&~~ + (1 - \gamma_{m}) u_{m+1}\\
&\leq f(m+1, p) + \sum_{k=1}^{m+1} \gamma_k' u_k
\end{align*}
\begin{align*}
&~~f_{\vee}(A^{(m+1)}, w) = f_{\vee}(A^{(m)}, w) + \gamma_{m+1}\prod_{k=1}^m(1 - w_k) w_{m+1}\\
&= f_{\vee}(A^{(m)}, w) +  \gamma_{m+1} \prod_{k=1}^m(1 - p_k) (p_{m+1} + u_{m+1})
\end{align*}
\end{proof}

@@

\begin{lemma}
Suppose $A = (a_1, ..., a_{\abs{A}})$. Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. Then we have the following properties.
\begin{itemize}
\item[(1)] $f(A, w) \leq p_A + \gamma_{\abs{A} - 1}$;
\item[(2)] $f(B_{k+1}, w) \leq f(B_k, w)$;
\item[(3)] $p_{B_{k+1}} \leq p_{B_k}$;
\item[(4)] $f(B_k, w) \leq p_{B_{k+1}} + \gamma_{k}$.
\end{itemize}
It also holds that $p_{A} \geq \gamma_{\abs{A}} - f_{\vee}(A, w)$.
\end{lemma}
